### 2023-06-08 paper/localpsf.tex
Editing based on discussion with Noemi (everything before numerical results)

% REQUIRED
\begin{abstract}
	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
%	, with a focus on approximating high rank Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs). 
%	The method may be used to approximate other operators, including certain Schur complements, Poincare-Steklov operators, covariance operators, and blurring operators. 
	The method computes impulse responses of the operator at a collection of scattered points, and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel. We compute a large number of impulse responses by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources. By solving an ellipsoid packing problem, we compute as many impulse responses as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap. Support ellipsoids are estimated a-priori based on impulse response moments, which are computed by applying the operator to a small number of polynomial functions. The interpolation uses a ``local mean displacement invariance'' approximation, which generalizes classical local translation invariance. The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator. Further matrix computations are performed with H-matrix methods. We apply the method to approximate the Hessians in an ice sheet flow inverse problem governed by a Stokes PDE, and a diffusive transport inverse problem governed by an advection-diffusion PDE. Numerical results demonstrate that the proposed method outperforms existing methods, and can approximate high rank operators using only a small number of operator applies.
\end{abstract}

>>>>

% REQUIRED
\begin{abstract}
	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
	The method computes impulse responses of the operator at a collection of scattered points, and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel. By applying the operator to Dirac combs associated with batches of point sources, we compute a large number of impulse responses using a small number of operator applies. We solve an ellipsoid packing problem to select as many point sources as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap. Support ellipsoids are estimated a-priori based on impulse response moments, which are computed by applying the operator to a small number of polynomial functions. The interpolation uses a local mean displacement invariance approximation, which generalizes classical local translation invariance. The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator. Further matrix computations are performed with H-matrix methods. We illustrate the method by approximating the Hessians in two example problems: an ice sheet flow inverse problem governed by a Stokes partial differential equation (PDE), and a diffusive transport inverse problem governed by an advection-diffusion PDE. Because the data in these inverse problems are highly informative about the unknown parameters, the data misfit terms in the Hessians are high rank. This makes conventional matrix-free Hessian approximation methods, which are based on low rank approximation, extremely expensive in terms of the required number of Hessian operator applies. In contrast, the numerical results show that the proposed method can form good approximations of these high rank Hessians using a small number of Hessian operator applies. We use the Hessian approximations as preconditioners to accelerate the solution of the inverse problems. Numerical results demonstrate that this approach substantially reduces the number of PDE solves required to solve the inverse problems.

%	, as compared to existing methods.

%	, as measured by the total number of PDE solves required to solve the inverse problem.

%	, with a focus on approximating high rank Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs). 
%	The method may be used to approximate other operators, including certain Schur complements, Poincare-Steklov operators, covariance operators, and blurring operators.

%	that arise in inverse problems with highly informative data.

%	we see that the numerical results show that the proposed method can approximate high Existing methods typically perform poorly  when the data is highly informative about the unknown parameter, because the Hessians in these problems are high rank. In contrast, the proposed method can cheaply approximate high rank Hessians.

%	, in the sense that it can approximate Hessians in inverse problem with highly informative data, using only a small number of operator applies. outperforms (as measured by number of PDE solves) existing methods for Hessian approximation. In particular, the proposed method can approximate Hessians in inverse problem with highly informative data, using only a small number of operator applies.
\end{abstract}
